How is experience transformed into memory?

### 2018-09-03 paper/main.tex
checking in changes after edits to wordles figure and associated text

Because our analysis framework projects the dynamic video content and participants' recalls onto a shared topic space, and because the dimensions of that space are known (i.e., each topic dimension is a set of weights over words in the vocabulary; Fig.~\topics), we can examine the topic trajectories to understand which specific content was remembered well (or poorly).  For each video event, we can ask: for all participants who recalled that event, what was the average correlation (across participants) between the video event's topic vector and the recall event topic vectors from each participant?  This yields a single correlation coefficient for each video event, describing how closely participants' recalls of the event tended to reliably capture its content (Fig.~\ref{fig:topics}A).  (We also examined how different comparisons between each video event's topic vector and the corresponding recall event topic vectors related to hand-annotated characterizations of memory performance; see \textit{Supplemental Materials}).  Given this summary of which events were recalled reliably (or not), we next asked whether the better-remembered or worse-remembered events tended to reflect particular topics.  We computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Events that reflected topics weighting heavily on characters like ``Sherlock'' and ``John'' (i.e., the main characters) and locations like ``221b Baker Street'' (i.e., a major recurring location; the address of the flat that Sherlock and John share) were best remembered.  An analogous analysis revealed which themes were poorly remembered; here in computing the weighted average over events' topic vectors we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  This revealed that events with relatively minor characters such as ``Mike,'' ``Jeffrey,'' and ``Molly,'' as well as less-integral plot locations (e.g., ``hospital'' and ``office'') were least well-remembered.  This suggests that what is retained in memory are the major plot elements (i.e., the overall ``gist'' of what happened), whereas the more minor details are prone to pruning.

In addition to constructing overall summaries, assessing the video and recall topic vectors from individual recalls can provide further insights.  Specifically, for any given event we can construct a wordle from that event's topic vector, and we can construct a similar wordle from the average topic vectors produced by all participants who recalled that event.  We can then examine those wordles visually to gain an intuition for which aspects of the video event were recapitulated in participants' recalls of that event.  Several example wordles are displayed in Figure~\ref{fig:topics}C (wordles from the three best-remembered events are circled in green; wordles from the three worst-remembered events are circled in red).  \textbf{[JRM NOTE: NEED SOME SORT OF ``POINT'' OR TAKE-AWAY FROM THIS ANALYSIS]}

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/topics}
\caption{\small \textbf{Transforming experience into memory.} \textbf{A.} Average correlations (across participants) between the topic vectors from each video event and the closest-matching recall events.  Error bars denote bootstrap-derived across-participant 95\% confidence intervals.  The stars denote the three best-remembered events (green) and worst-remembered events (red).  \textbf{B.} Wordles comprising the top 200 highest-weighted words reflected in the weighted-average topic vector across video events.  Green: video events were weighted by how well the topic vectors derived from recalls of those events matched the video events' topic vectors (Panel A).  Red: video events were weighted by the inverse of how well their topic vectors matched the recalled topic vectors.  \textbf{C.}  The set of all video and recall events is projected onto the two-dimensional space derived in Figure~\ref{fig:trajectory}.  The large dots denote video events and the smaller dots denote recalled events (same color scheme as Fig.~\ref{fig:trajectory}A).  Wordles for several example events are displayed (green: three best-remembered events; red: three worst-remembered events).  Within each circular wordle, the left side displays words associated with the topic vector for the video event, and the right side displays words associated with the (average) recall event topic vector, across all recall events matched to the given video event.}
\label{fig:topics}
\end{figure}

>>>>

Because our analysis framework projects the dynamic video content and participants' recalls onto a shared topic space, and because the dimensions of that space are known (i.e., each topic dimension is a set of weights over words in the vocabulary; Fig.~\topics), we can examine the topic trajectories to understand which specific content was remembered well (or poorly).  For each video event, we can ask: what was the average correlation (across participants) between the video event's topic vector and the closest matching recall event topic vectors from each participant? This yields a single correlation coefficient for each video event, describing how closely participants' recalls of the event tended to reliably capture its content (Fig.~\ref{fig:topics}A).  (We also examined how different comparisons between each video event's topic vector and the corresponding recall event topic vectors related to hand-annotated characterizations of memory performance; see \textit{Supplemental Materials}).  Given this summary of which events were recalled reliably (or not), we next asked whether the better-remembered or worse-remembered events tended to reflect particular topics.  We computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Events that reflected topics weighting heavily on characters like ``Sherlock'' and ``John'' (i.e., the main characters) and locations like ``221b Baker Street'' (i.e., a major recurring location; the address of the flat that Sherlock and John share) were best remembered.  An analogous analysis revealed which themes were poorly remembered; here in computing the weighted average over events' topic vectors we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  This revealed that events with relatively minor characters such as ``Mike,'' ``Jeffrey,'' and ``Molly,'' as well as less-integral plot locations (e.g., ``hospital'' and ``office'') were least well-remembered.  This suggests that what is retained in memory are the major plot elements (i.e., the overall ``gist'' of what happened), whereas the more minor details are prone to pruning.

In addition to constructing overall summaries, assessing the video and recall topic vectors from individual recalls can provide further insights.  Specifically, for any given event we can construct a wordle from that event's topic vector, and we can construct a similar wordle from the average topic vectors produced by all participants who recalled that event.  We can then examine those wordles visually to gain an intuition for which aspects of the video event were recapitulated in participants' recalls of that event.  Several example wordles are displayed in Figure~\ref{fig:topics}C (wordles from the three best-remembered events are circled in green; wordles from the three worst-remembered events are circled in red).  Using wordles to visually compare the topical content of each video event and the (average) corresponding recall event reveals the specific content from the specific events that is reliably retained in the transformation into memory (green events) or not (red events).

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/topics}
\caption{\small \textbf{Transforming experience into memory.} \textbf{A.} Average correlations (across participants) between the topic vectors from each video event and the closest-matching recall events.  Error bars denote bootstrap-derived across-participant 95\% confidence intervals.  The stars denote the three best-remembered events (green) and worst-remembered events (red).  \textbf{B.} Wordles comprising the top 200 highest-weighted words reflected in the weighted-average topic vector across video events.  Green: video events were weighted by how well the topic vectors derived from recalls of those events matched the video events' topic vectors (Panel A).  Red: video events were weighted by the inverse of how well their topic vectors matched the recalled topic vectors.  \textbf{C.}  The set of all video and recall events is projected onto the two-dimensional space derived in Figure~\ref{fig:trajectory}.  The dots outlined in black denote video events (dot size reflects the average correlation between the video event's topic vector and the topic vectors from the closest matching recalled events from each participant; bigger dots denote stronger correlations).  The dots without black outlines denote recalled events.  All dots are colored using the same scheme as Figure~\ref{fig:trajectory}A.  Wordles for several example events are displayed (green: three best-remembered events; red: three worst-remembered events).  Within each circular wordle, the left side displays words associated with the topic vector for the video event, and the right side displays words associated with the (average) recall event topic vector, across all recall events matched to the given video event.}
\label{fig:topics}
\end{figure}
